Distributed Web Mining of Ethereum

### 2018-04-09 paper/paper.tex
almossst done

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{color}
\usepackage{makecell}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{amssymb}
\newcommand{\trishita}[1]{{\color{magenta}\bfseries[Trishita: #1]}}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\begin{document}
%
\title{Distributed Web Mining of Ethereum\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Trishita Tiwari%\orcidID{0000-1111-2222-3333} 
\and
David Starobinski%\orcidID{1111-2222-3333-4444} 
\and
Ari Trachtenberg%\orcidID{2222--3333-4444-5555}
}
%
\authorrunning{T. Tiwari et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Boston University, Department of Electrical and Computer Engineering, Boston, MA 02215, USA \\
\email{\{trtiwari,staro,trachten\}@bu.edu}}
%
\maketitle             
%
\begin{abstract}
We consider the problem of mining crytocurrencies by harnessing the inherent distribution capabilities of the World Wide Web. More specifically, we consider
the mining of the Ethereum cryptocurrency by apportioning elements of the corresponding Proof-of-Work calculations among a number of web clients. These calculations are handled through individualized code that runs on the client browsers, and is thereafter collated at a web server to complete a mining operation. Our approach is based on a lazy evaluation technique designed to function within the expected limitations of the clients, including bounds on memory, computation and communication bandwidth to the server. We provide proofs-of-concept based in JavaScript and WebAssembly, with the latter reaching hash rates up to roughly 39 kiloHashes per second, which is only 30.2\% slower than the corresponding native C++-based miner.
Finally, we explore several applications for which this hash rate seems practical, including monetization of web content, rate limitation, and private Ethereum networks. Though several distributed web-based cryptominers have appeared in the wild (for other currencies), either in malware or in commercial trials, we believe that this is the first open-source cryptominer of this type.
% The process of legitimating cryptocurrencies (such as Bitcoin and Ethereum) is done
% through mining, often in a competitive and distributed manner by means of a small number of
% large, loosely-organized mining colonies following proprietary protocols.  We

@@

% Given the seemingly endless growing popularity of crypto-currencies, there is an increasing interest in different mining possibilities. This paper proposes a web browser based mining implementation for Ethereum implemented both in JavaScript and WebAssembly to perform the necessary Proof-of-Work (PoW) computations to mine the Ethereum blockchain. We explore a lazy evaluation based technique to distribute the inherently memory and memory intensive PoW of Ethereum over multiple browser-based clients that have limited resources. We obtain the best results with the WebAssembly implementation, with hash rates reaching upto 39651H/s ~\trishita{Not sure if we need such a specific number}, which, as we demonstrate, is merely 30.2\% slower than a native miner that uses the same algorithm for mining under the same conditions. We show that this solution is practical for a variety of purposes, such as web content monetization, web based authentication rate limiting and private Ethereum networks.
% \trishita{I have not included a comparison with an actual miner that saves the entire DAG in memory from the start (i.e., doesn't use our lazy evaluation approach) -- should I do that for the abstract?}

\keywords{crypto-currency \and Ethereum \and distributed computing \and web-browser computing \and mining}
\end{abstract}
%
%
%
\section{Introduction}
Cryptocurrencies are increasingly gaining traction as a viable form of currency.  This has been accompanied by a correspondingly increasing interest in the efficient validation (or \emph{mining}) of cryptocurrency transactions.  Whereas initial efforts in this domain have focused on creating dedicated hardware for this task, %AT - citation needed
more recent approaches have examined repurposing existing infrastructure.  Indeed, one such class of efforts has focused on the use of client web browsers as a platform for distributed computing~\cite{Cushing}. The growing popularity of CoinHive~\cite{coinhive} is a case in point of the potential success of distributed in-browser cryptocurrency mining as a commercial (if malicious) enterprise.

\subsection{Our Contributions}
In this work, we explore the web-based crypocurrency mining of the popular Ethereum cryptocurrency, %AT - citation needed
and posit a number of potential applications consequent to this platform.  As such, our main contributions are as follows:
\begin{itemize}
\item We provide an open-source implementaiton of a web-based Ethereum cryptominer in both \verb|JavaScript| and \verb|WebAssembly|.  Though miners for other currencies exist in the wild (e.g. Coinhive), they are all properietary and closed-source.
\item We posit a number of potential applications built upon such a platform, including rate-limiting server access, usage tracking, and content monetization.
\item We provide analysis and experimental evidence of the potential efficacy of our approach to the applications above.
\end{itemize}

@@

\subsection{Roadmap}
The remainder of the paper is organized as follows. In Section~\ref{sec:applications} we discuss various potential applications our mining platform. Then, in Section~\ref{sec:background} we cover the relevant background and related literature for our work.  In particular, we provide a self-contained description of Ethash, the Ethereum mining Proof-of-Work (PoW), and \verb|WebAssembly|, the language in which we implement an efficient miner. Section~\ref{sec:approach} outlines our approach to distributing the Ethash PoW over numerous browsers, each with their own resource constraints. Finally, we present the results in Section~\ref{sec:results}. We conclude and discuss the future work in Sectionn~\ref{sec:conclusion}.

\section{Applications}
\label{sec:applications}
% In this work, we propose approaches to mine Ethereum, a popular crypto-currency, in a distributed fashion through web browsers. Our approach can be used for a variety of purposes, outlined as follows. \\ \\
% \textbf{Web Content Monetization}
With the growth in global Internet usage, hosting websites has become a lucrative business. As a result, new methods of monetizing electronic content have surfaced with time. While some are more successful than others, all of them have issues associated with them. For instance, selling advertisement space is now resulting in declining revenue \cite{decliningRevenue} for website owners due to the advent of new technologies such as AdBlock \cite{Adblock}, Brave Browser \cite{BraveBrowser}, etc. In addition, too many advertisements also tend to increase the loading time for web-pages. Moreover, such advertisements are sometimes not placed correctly and cover the contents of the website. Thus, this approach does not provide good user experience. \\ \\
% Another approach involves subscriptions, however, it is not very successful. Indeed, as this study finds, users are becoming increasingly unlikely to pay for content \cite{noSubscriptions}. Website owners also use pay-per-click Advertising, however an approach often leads to the content being split over an excessive number of pages, thereby leading to consumer frustration \cite{payPerClick}. Thus, we see that each approach has its own issues, be it with  with website owners or clients. Hence, it is important to strike a balance and create a solution that would be able to benefit both parties. Our web-based miner would solve this situation as it does not 
% \textbf{Web-Authentication Rate Limiting}: 
Another potential application could involve web-authentication rate limiting. Given the multitude of tools available brute forcing web login pages ~\cite{hydra} ~\cite{burpsuite}, it is extremely easy to brute force authentication pages these days. Currently, the way website owners mitigate this issue is by locking out a user for a certain amount of time after a fixed number of unsuccessful login attempts. 
% While this approach works in theory, in today's fast-paced world, users tend to dislike waiting for even a few seconds ~\cite{usersDislikeWaiting}. 
While it seems as though this approach is effective at face value, in reality, it does more harm than good as by introducing the risk of Denial of Service. This is because most brute force attacks are large scale and try millions of username/password combinations, which would consequently lockout a large number of user accounts. While this may prevent an attacker from successfully cracking passwords, it would also deny service to a lot of legitimate users whose accounts were being brute forced. This, amongst other issues, would definitely reflect badly on the website owner, and hence opens a whole new avenue for attack.   
Given this situation, a more user-friendly approach to this problem involves embedding a Proof-of-Work (PoW) computation in the web page that the user's browser needs to successfully solve in order to be able to login. Requiring a PoW would require a large amount of computation power for brute force attempts, thereby selectively thwarting any attacker that attempts to brute force the login \emph{without} penalizing the legitimate user. This makes our implementation for the Ethereum PoW an ideal candidate, even more so since one could manually set the difficulty to obtain a balance between user experience and security. \\ \\
One other use-case could involve website advertisement companies. Today, website advertising sponsors decide on the remuneration for a website by inspecting the logs of the web servers in order to deduce how much traffic the site receives. However, it is well known that these logs can be easily manipulated by the website owner in order to generate the impression of a large amount of traffic ~\cite{webLogsManipulation}. Our web miner could solve these issues as the advertiser could require the website owner to embed this miner within their website, thereby making the site visitors compute PoW hashes. The advertiser would then ask the website owner to submit hashes that pass a certain difficulty threshold (i.e., the value of the hash being less than a certain number), and the larger the number of hashes that the website owner can provide, the more the remuneration the site receives. This would be extremely difficult for the website owners to fake since they would have to compute hashes themselves, an endeavor that might be more expensive than the corresponding payout it would bring.\\ \\
Finally, we would also like to note that Ethereum is an extremely flexible currency in the sense that it allows for private coin networks -- i.e., networks that do not mine on the actual, public Ethereum block-chain, but rather on private (and often smaller) instances of the crypto-currency. Our browser miner can be used on any such private network to serve the network owner whatever he/she wishes to implement with the mining PoW.

\section{Background}
\label{sec:background}
% \subsection{Ethereum}
% \trishita{Not sure if I need this paragraph, since I think the readers will probably be familiar with what crypto-curriencies are}
\subsection{JavaScript and WebAssembly}
For very long, \verb|JavaScript| has been the only language available to make dynamic client side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. As per its creators, ``\verb|WebAssembly| is a binary instruction format for a stack-based virtual machine. Wasm is designed as a portable target for compilation of high-level languages like \verb|C/C++/Rust|, enabling deployment on the web for client and server applications. The \verb|WebAssembly| stack machine is designed to be encoded in a size- and load-time-efficient binary format. \verb|WebAssembly| aims to execute at native speed by taking advantage of common hardware capabilities available on a wide range of platforms'' ~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc ~\cite{webAssembly}. This, and given the fact that CoinHive is also implemented in \verb|WebAssembly| ~\cite{coinhive}, makes the language a natural choice for us to implement a miner in. In fact, as of now, \verb|WebAssembly| is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}, which makes our miner implementation extremely versatile.

% Like most crypto-currencies, it too is backed by a distributed block-chain, each block of which holds transactions for that particular block-chain. This block chain is operated on by the Ethereum Virtual Machine (EVM) ~\cite{EVM}, which, unlike Bitcoin Script, is Turing Complete ~\cite{turingComplete}. Furthermore, this currency is also a Proof-of-Work based currency, which means that the new blocks are appended to the block-chain through a process called ``mining''. Mining involves a race amongst miners to solve a Proof-of-Work puzzle (usually an energy intensive computation), the winner of which gets to have his block appended to the block-chain. The winner also receives a payout, which acts as an incentive to mine. This process of mining is essential to deciding the order in which transactions are appended to the inherently decentralized block-chain, thereby stabilizing the chain. Ethereum uses Ethash as its Proof-of-Work (PoW) algorithm, which is explained in detail below.

\subsection{Ethash}
Ethereum is a crypto-currency that was released in July 2015 by Vitalik Buterin, Gavin Wood and Jeffrey Wilcke. Ethereum uses the Ethash algorithm (derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}) for its PoW for mining blocks. Before we discuss how mining works with Ethash, we would first like to establish some basic terminology about the data structures involved in the PoW:
\begin{itemize}
\item \textbf{Seed}: Ethash operates by first generating a 256 bit seed. This seed is different for every epoch (30,000 blocks). For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash. 
\item \textbf{Cache}: The seed for a particular epoch is used to compute a pseudo-random cache. This cache is simply an array 4 byte integers ~\cite{Ethmining}. The cache production process involves using the seed hash to first sequentially fill up the Cache, then performing two passes of the RandMemoHash algorithm~\cite{randmemohash} on top of it to get the final Cache value. 
% This transforms the initial cache to give us the final value for the cache. 
Lightweight clients may use this cache for verifying hashes (discussed below), while full node clients can use it to calculate a DAG (Directed Acyclic Graph) dataset, as described below. 
\item \textbf{DAG}: A Directed Acyclic Graph stored as a large byte array (currently around 2GB in size). The DAG has the following two attributes:
  \begin{enumerate}
  \item \textbf{Node}: Each DAG node in this byte array spans 64 bytes, and node indices are therefore aligned at a 64 byte boundary.
  \item \textbf{Page}: Each DAG page spans 2 nodes, however, page accesses are not aligned at a 2 node boundary. The mining process involves accessing some DAG pages and hashing them together with the header and nonce.
  \end{enumerate}
Each node in the DAG is generated by combining data from 256 pseudorandomly selected cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG.
\end{itemize}
We must point out that as time goes on, mining Ethereum becomes more and more difficult, as the size of the DAG and Cache increases with every Epoch.

@@

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array),as seen in step 1 of Figure~\ref{fig:ethash}. The Mix is then used to fetch a specific page of the DAG from memory. After this, the Mix is updated with the fetched part of the DAG (step 2). Then, this updated Mix is used to fetch a new part of the DAG (step 3). This process of sequentially fetching parts of the DAG is repeated 64 times (step 4), and the final value of the Mix is put through a transformation to obtain a 32 byte digest (step 5). This digest is then compared to the threshold (step 6), and if it is lesser than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce is unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}. Below is the pseudo-code for the algorithm for completeness.\\
\begin{algorithmic}[H]
\label{miningAlgo}
\STATE $mix \gets mixInit(\{header,nonce\})$
\FOR{$i = 0;\;i < 64;\;i++$}
  \STATE $dagPageIndex \gets getDAGPageIndexFromMix(mix)$
  \STATE $dagPage \gets DAG[dagPageIndex]$
  \STATE $dagPageIndex \gets updateMix(dagPage,mix)$
\ENDFOR
\STATE $digest \gets postProcessFunction(mix)$
\IF {$digest <= targetThreshold$}
	\RETURN \TRUE
\ENDIF
\RETURN \FALSE
\end{algorithmic}
It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence there is no way to pre-determine which parts will be useful to have in memory. This therefore forces miners to store entire DAG in memory, making mining ``Memory Hard''.

@@

While mining is memory intensive, verification is relatively lightweight. This is because of the property that each ``slice'' in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific slices of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by clients that perform just verification. We use this property of being able to generate the parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

@@

\subsection{Challenges}
% \trishita{Too long of a sentence; need to reword}
Given that we now know that the crux of having a good hashrate for Ethereum lies in having quick access to the DAG, one of the biggest challenges with mining it on a distributed ensemble of browsers would be to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 1GB (and growing) in size transferring it to the browser every time a client loads our web page would be impractical. Furthermore, even if we somehow manage to transfer the DAG to the client when the web page loads, the browser would still have to store it in memory, thereby using a sizeable amount of resources on the client machine. We attempt to address some of these concerns through our approach as discussed in the next section.

>>>>

% 
% instead of saying lazy evaluation -- say how much time do we take 
% some quantification on how much time does it take to get 
% put challenges in introduction -- what does this make hard to solve? Put in the introduction. 
% intro, one line on contributions, and put challenges in the beginning. -- this will allow to motivate better (can use dag, say you will explaon)

% \para{application}: then application. Then a new para how our approach can be applied to the problem. Put the whole section at the end. Also put the references.

% Don't use quotes on very long sentences. 
% Put Javascrip Wasm section after Ethash. This section is too small. Do we really want it. 
% Talk about wasm in one place, not two.

% remove pseudocode.
% Have a section of ethereum and define what block header and hash. Also define target etc 
% dEFINE epoch early on. 
% called Cache in ethereum terminology. Mention that the size of the cache varies right 
% define lightweight and full node earlier.
% Say each DAG consists of nodes and slices. Explain what 
% Ads: talk about how our approach can be applied to

% our approach --
% how long -- our key contribution is to measure if this is viable to wait till the hash rate gets good. We need to mention this earlier. 
% remove `n'
% the video application. 
% Have a name. 
% its not clear that js was used. 
% have a para on js and webasm in our approach. 
% have the definitions before the claim. 
% use a different var for Hash and 
% put in the logs!!

% look at all citations -- at least there should be a year and conference. 
% more descriptive captions for figures.

% minimize the jargon
% experiments should reinforce what you say before.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{color}
\usepackage{makecell}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{caption}
\newcommand{\trishita}[1]{}%{{\color{magenta}\bfseries[Trishita: #1]}}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\begin{document}
%
\title{Distributed Web Mining of Ethereum
% \thanks{Supported by organization x.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Trishita Tiwari%\orcidID{0000-1111-2222-3333} 
\and
David Starobinski%\orcidID{1111-2222-3333-4444} 
\and
Ari Trachtenberg%\orcidID{2222--3333-4444-5555}
}
%
\authorrunning{T. Tiwari et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Boston University, Department of Electrical and Computer Engineering, Boston, MA 02215, USA \\
\email{\{trtiwari,staro,trachten\}@bu.edu}}
%
\maketitle             
%
\begin{abstract}
We consider the problem of mining crytocurrencies by harnessing the inherent distribution capabilities of the World Wide Web. More specifically, we consider
the mining of the Ethereum cryptocurrency by apportioning elements of the corresponding Proof-of-Work calculations among a number of web clients. These calculations are handled through individualized code that runs on the client browsers, and is thereafter collated at a web server to complete a mining operation. WebEth is based on a lazy evaluation technique designed to function within the expected limitations of the clients, including bounds on memory, computation and communication bandwidth to the server. We provide proofs-of-concept based in JavaScript and WebAssembly, with the latter reaching hash rates up to roughly 40 kiloHashes per second, which is only 30\% slower than the corresponding native C++-based miner.
Finally, we explore several applications for which this hash rate seems practical, including monetization of web content, rate limitation, and private Ethereum networks. Though several distributed web-based cryptominers have appeared in the wild (for other currencies), either in malware or in commercial trials, we believe that this is the first open-source cryptominer of this type.
% The process of legitimating cryptocurrencies (such as Bitcoin and Ethereum) is done
% through mining, often in a competitive and distributed manner by means of a small number of
% large, loosely-organized mining colonies following proprietary protocols.  We

@@

% Given the seemingly endless growing popularity of crypto-currencies, there is an increasing interest in different mining possibilities. This paper proposes a web browser based mining implementation for Ethereum implemented both in JavaScript and WebAssembly to perform the necessary Proof-of-Work (PoW) computations to mine the Ethereum blockchain. We explore a lazy evaluation based technique to distribute the inherently memory and memory intensive PoW of Ethereum over multiple browser-based clients that have limited resources. We obtain the best results with the WebAssembly implementation, with hash rates reaching upto 39651H/s ~\trishita{Not sure if we need such a specific number}, which, as we demonstrate, is merely 30.2\% slower than a native miner that uses the same algorithm for mining under the same conditions. We show that this solution is practical for a variety of purposes, such as web content monetization, web based authentication rate limiting and private Ethereum networks.

\keywords{crypto-currency \and Ethereum \and distributed computing \and web-browser computing \and mining}
\end{abstract}
%
%
%
\section{Introduction}
Cryptocurrencies are increasingly gaining traction as a viable form of currency.  This has been accompanied by a correspondingly increasing interest in the efficient validation (or \emph{mining}) of cryptocurrency transactions.  Whereas initial efforts in this domain have focused on creating dedicated hardware for this task, %AT - citation needed
\trishita{Add citation}
more recent approaches have examined repurposing existing infrastructure. Indeed, one such class of efforts has focused on the use of client web browsers as a platform for distributed computing~\cite{Cushing}. The growing popularity of CoinHive~\cite{coinhive} is a case in point of the potential success of distributed in-browser cryptocurrency mining as a commercial (if malicious) enterprise. In this work, we propose WebEth, a browser-based distributed miner for the Ethereum block chain. Obtaining a profitable hash rate within a distributed ensemble of browsers under constrained memory, compute and network usage is non-trivial. This is primarily because every worker needs to store a 2GB data structure in memory in order to mine Ethereum. However, the naive approach of transferring the data structure over the network and storing it in memory every time a browser loads our web page is not viable. And so, WebEth employs an amortized approach to mining in order to operate under such constraints. In our experiments, this approach takes at most 5 minutes to reach a steady state hash rate -- making it ideal for websites where users spend more time, like video streaming websites. WebEth yields at most 40 kiloHashes/s, which despite the overhead from running the algorithm in a browser, is only 30\% lesser than the performance of the corresponding native miner that employs our algorithm, thereby proving that our approach is effective.

\subsection{Our Contributions}
In this work, we explore the web-based cryptocurrency mining of the popular Ethereum block chain, \trishita{Add citation}
%AT - citation needed
and propose a number of potential applications consequent to this platform.  As such, our main contributions are as follows:
\begin{itemize}
\item We propose WebEth, an open-source implementation of a distributed web-based Ethereum cryptominer in both \verb|JavaScript| and \verb|WebAssembly| that can operate under relatively resource constrained environments. Though miners for other currencies exist in the wild (e.g. CoinHive~\cite{coinhive}), they are all proprietary and closed-source.
\item We propose a number of potential applications built upon such a platform, including rate-limiting server access, usage tracking, and content monetization.
\item We provide analysis and experimental evidence of the potential efficacy of our approach to the applications above. \trishita{I feel like this next part is repetitive}Our miner employs an amortized approach to mining that takes $\approx$ 5 minutes for it to reach its maximum hash rate. In addition, we show that our \verb|WebAssembly| implementation can provide hash rates up to 40 kiloHashes/s, which is only 30\% smaller than the hash rate provided by a native C++ miner that employs the same algorithm. 
\end{itemize}

@@

\subsection{Roadmap}
\trishita{Fix this acc to the new orderigns}
The remainder of the paper is organized as follows. In Section~\ref{sec:background} we cover the relevant background and related literature for our work.  In particular, we provide a self-contained description of Ethash, the Ethereum mining Proof-of-Work (PoW), and \verb|WebAssembly|, the language in which we implement an efficient miner. Section~\ref{sec:approach} outlines WebEth's approach to distributing the Ethash PoW over numerous, resource-constrained browsers. Then, we present the results in Section~\ref{sec:results}. Finally, in Section~\ref{sec:applications} we discuss various potential applications our mining platform. We conclude and discuss the future work in Sectionn~\ref{sec:conclusion}.

\section{Background}
\label{sec:background}
\subsection{Cryptocurrencies: a general overview}
% \trishita{Not sure if I need this paragraph, since I think the readers will probably be familiar with what crypto-curriencies are}

Most cryptocurrencies like Ethereum involve storing transactions in blocks, and the entire history of transactions is collated in a data structure known as a block chain. The block chain is managed by dedicated machines called ``client nodes''; where each client node typically stores the entire block chain. However, because each client node operates on its own copy of the block chain, it is very easy for the block chains on different nodes to go out of sync. Hence, there needs to be an accepted mechanism to decide the order in which new transaction blocks are appended to this block chain. 
For this, every time a new transaction takes place, the transaction is thrown in with all other transactions that have ``taken place'', but haven't been added to the block chain yet. For most currencies, the data structure that stores this is known as the ``mem pool''. A miner then picks a few transactions from this mem pool and creates a new block from these transactions. Once this happens, the goal of the miner is to have his/her block appended to the block chain, which is achieved through a process called ``mining''.
% This block chain is operated on by the Ethereum Virtual Machine (EVM) ~\cite{EVM}, which, unlike Bitcoin Script, is Turing Complete ~\cite{turingComplete}. 
% Furthermore, this currency is also a Proof-of-Work based currency, which means that the new blocks are appended to the block-chain through a process called ``mining''. 
Mining involves a race amongst miners to solve a ``Proof-of-Work puzzle'' (usually an energy intensive computation), the winner of which gets to have his block appended to the block-chain. The winner also receives a payout, which acts as an incentive to mine. 
% This process of mining is essential to deciding the order in which transactions are appended to the inherently decentralized block-chain, thereby stabilizing the chain. 
Ethereum uses Ethash as its Proof-of-Work (PoW) algorithm, which is explained in detail below.

\subsection{Ethereum and its PoW, Ethash}
Ethereum is a crypto-currency that was released in July 2015 by Vitalik Buterin, Gavin Wood and Jeffrey Wilcke. Ethereum uses the Ethash algorithm (derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}) for its PoW for mining blocks. Before we discuss how mining works with Ethash, we would first like to establish some basic terminology about the data structures involved in the PoW:
\begin{itemize}
\item \textbf{Block Header}: A block header contains meta-data related to the transactions of the block it pertains to, and is provided as an input to the Ethash algorithm.
\item \textbf{Nonce}: An integer that is provided as an input to the hashing algorithm in the hopes that when hashed with the block header, will yield a hash value less than a certain predefined difficulty threshold. It is also provided as an input to Ethash.
\item \textbf{Difficulty Threshold}: An integer value below which the hash digest output of Ethash should be in order for the miner to have successfully mined the block.
\item \textbf{Mining overview}: The Ethash mining algorithm primarily involves computing a hash of the current block header and nonce. If the output of the hash (the digest) lies below the difficulty threshold; the block is said to have been successfully mined. Hence, the miner keeps iterating over nonces until he/she gets a digest that lies below the difficulty threshold. Once the miner finds the right nonce, he/she then broadcasts the block, the header and the nonce so that the other client nodes can update their block chain to include this newly mined block. 
\item \textbf{Verification}: Once a block has been mined by a miner, it is propagated to other client nodes so that they can update their copies of the block chain. However, before each client node does so, it must validate whether the miner is submitting a legitimate block -- i.e., check whether the miner genuinely solved the hash as claimed. This is easily done by putting the block header and nonce associated with the block through the Ethash algorithm and checking whether the output indeed lies below the difficulty threshold. 
\item \textbf{Light Weight Client Nodes}
Some client nodes do not mine new blocks, but just verify whether any new block submitted by a miner is valid or not. These are called light weight client nodes.
\item \textbf{Full Client Nodes}
Full client nodes mine new blocks.
\item \textbf{Epoch}: A unit of ``time'' that spans 30,000 blocks.
\item \textbf{Seed}: All of the data structures in Ethash (mentioned below) need a 256 bit seed. This seed is different for every Epoch. For the first epoch it is the Keccak-256~\trishita{CITE} hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash. 
\item \textbf{Cache}: The seed for a particular epoch is used to compute a data structure known as the Cache. This Cache is simply an array 4 byte integers ~\cite{Ethmining}. The cache production process involves using the seed hash to first sequentially fill up the Cache, then performing two passes of the RandMemoHash algorithm~\cite{randmemohash} on top of it to get the final Cache value. 
% This transforms the initial cache to give us the final value for the cache. 
Light weight clients may use this cache for verifying hashes (discussed below), while full node clients can use it to calculate a Directed Acyclic Graph (DAG) dataset, as described below. 
\item \textbf{DAG}: A Directed Acyclic Graph stored as a large byte array (currently around 2GB in size). The DAG has the following two attributes:
  \begin{enumerate}
  \item \textbf{Node}: Each DAG node in this byte array spans 64 bytes, and node indices are therefore aligned at a 64 byte boundary.
  \item \textbf{Page}: Each DAG page spans 2 nodes, however, page accesses are not aligned at a 2 node boundary. The mining process involves accessing some DAG pages and hashing them together with the block header and nonce.
  \end{enumerate}
Each node in the DAG is generated by combining data from 256 pseudo-randomly selected cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG.
\item \textbf{Mix:} An intermediate byte array used to store temproary results in the Ethash algorithm.
\end{itemize}
We must point out that as time goes on, mining Ethereum becomes more and more difficult, as the size of the Cache and DAG increases with every Epoch.

@@

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array),as seen in step 1 of Figure~\ref{fig:ethash}. The Mix is then used to fetch a specific page of the DAG from memory. After this, the Mix is updated with the fetched part of the DAG (step 2). Then, this updated Mix is used to fetch a new part of the DAG (step 3). This process of sequentially fetching parts of the DAG is repeated 64 times (step 4), and the final value of the Mix is put through a transformation to obtain a 32 byte digest (step 5). This digest is then compared to the threshold (step 6), and if it is smaller than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce is unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}.

% Below is the pseudo-code for the algorithm for completeness.\\
% \begin{algorithmic}[H]
% \label{miningAlgo}
% \STATE $mix \gets mixInit(\{header,nonce\})$
% \FOR{$i = 0;\;i < 64;\;i++$}
%   \STATE $dagPageIndex \gets getDAGPageIndexFromMix(mix)$
%   \STATE $dagPage \gets DAG[dagPageIndex]$
%   \STATE $mix \gets updateMix(dagPage,mix)$
% \ENDFOR
% \STATE $digest \gets postProcessFunction(mix)$
% \IF {$digest <= targetThreshold$}
% 	\RETURN \TRUE
% \ENDIF
% \RETURN \FALSE
% \end{algorithmic}

It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence there is no way to pre-determine which parts will be useful to have in memory. This therefore forces miners to store entire DAG in memory, making mining ``Memory Hard''.

@@

While mining is memory intensive, verification is relatively lightweight. This is because of the property that each node in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific nodes of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by light weight clients that only perform verification. We use this property of being able to generate parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

@@

\subsection{JavaScript and WebAssembly}
Introduced in 1995 by Netscape Communications Corporation~\cite{https://auth0.com/blog/a-brief-history-of-javascript/}, \verb|JavaScript| was a meant to be a light scripting language in order to  make web content dynamic. Over the span of 23 years, it has grown to become one of the most popular client-side web development languages used to make dynamic user interfaces.

In fact, up to recently, \verb|JavaScript| has been the only language available to make dynamic client side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. As per its creators, \verb|WebAssembly| is a ``binary instruction format for a stack-based virtual machine''. WASM is designed to be compiled from high-level languages like \verb|C/C++/Rust|, and is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}. The \verb|WebAssembly| stack machine is designed to be encoded in a ``size-and load-time-efficient binary format''~\cite{webAssembly}, and aims to execute near native speed by utilizing common hardware capabilities available on a wide range of platforms~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc~\cite{webAssembly}. This, and given the fact that CoinHive is also implemented in \verb|WebAssembly| ~\cite{coinhive}, makes the language a natural choice for us to implement a miner in. 
% In fact, as of now, \verb|WebAssembly| is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}, which makes our miner implementation extremely versatile.

\subsection{Challenges}
% \trishita{Too long of a sentence; need to reword}
The crux of having a good hashrate for Ethereum lies in having quick access to the DAG. One of the biggest challenges with mining on a distributed ensemble of browsers is to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 1GB (and growing), transferring it to the browser every time a client loads our web page would be impractical. Furthermore, even if one somehow manages to transfer the DAG to the client when the web page loads, the browser still has to store the DAG in memory, thereby using a sizeable amount of resources on the client machine. In section~\ref{sec:approach} we discuss approaches to tackle some of these challenges.
